---
title: "CIS Project data-prep"
output: html_notebook
---

Create daatsets for Sherya based on the US maket. Thsi work was done on the 14 November 2021.

Focus on attrition at the OS level.

* Attrition data from 2019-2021.
* Active workforce at the end of every year - 2019-2021.
* Dataset with active and attrition for 2019-2021 with age profile, work experience etc - limited to OS.

The data for 2021 is until 30 Sep.

# Load the libraries:

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(readxl)
library(ggthemes)
library(tidytext)
library(ggridges)
theme_set(theme_economist_white())
```

# Load thee data:

```{r}
# promos
promo_2019 <- read_excel("data/Promotions_2019.xlsx") %>% clean_names()
promo_2020 <- read_excel("data/Promotions_2020.xlsx") %>% clean_names()
promo_2021 <- read_excel("data/Promotions_2021.xlsx") %>% clean_names()

promo_2020 <- promo_2020 %>% rename(gender = grade_2) #fix var name

promo_2019_2021 <- bind_rows(promo_2019, promo_2020, promo_2021) # joind all promos

# actives
active_2021 <- read_excel("data/WF Sep 2021_copy_pasted.xlsx", 
    sheet = "Workforce Sept") %>% clean_names()

active_2020 <- read_excel("data/WF report 2020-2021_03062021.xlsx", 
    sheet = "WF 2020") %>% clean_names()

active_2019 <- read_excel("data/WF report 2017-2019_03062021.xlsx", 
    sheet = "WF 2019") %>% clean_names()

# attritio

```

# Trim the dataset to just US [change ths to Finance]

```{r}

promo_amr_OS <- promo_2019_2021 %>% filter(region == "AMR",
                                            grade == "OS")

promo_amr_OS <- promo_amr_OS %>% 
  mutate(adjustment_date = parse_date_time(adjustment_date, orders = c("ymd")),
         year = year(adjustment_date))

attrition_amr_os <- attrition_2017_2021 %>% filter(region_code == "AMR",
                                                   leave_date > dmy("31-12-2018"),
                                                   cor_grade == "OS")

active_2019_amr_os <- active_2019 %>% filter(region_code == "AMR",
                                            cor_grade == "OS") %>% 
  mutate(data_year = 2019) # add to distinguish aactive data for each year
  
active_2020_amr_os <- active_2020 %>% filter(region_code == "AMR",
                                            cor_grade == "OS")%>% 
  mutate(data_year = 2020) # add to distinguish aactive data for each year

active_2021_amr_os <- active_2021 %>% filter(region_code == "AMR",
                                            cor_grade == "OS") %>% 
  mutate(data_year = 2021) # add to distinguish aactive data for each year
  
```

# Tables for Shreya
* 2019-2021, actives, attrition, promotions for OS tabulated form
* Combine the actives, promotios, attrition data for 2019-2021 for OS in raw form


```{r}
# Raw form data.
# select relevant cols

# actives:
active_data_combined <- bind_rows(
  active_2019_amr_os %>% 
    select(empl_no, employee_name, position_title, job_family, hire_date, serv_yrs, country, gender, birth_date, nationality, cor_grade, data_year),

  active_2020_amr_os %>% 
    select(empl_no, employee_name, position_title, job_family, hire_date, serv_yrs, country, gender, birth_date, nationality, cor_grade, data_year),

  active_2021_amr_os %>% 
    select(empl_no, employee_name, position_title, job_family, hire_date, serv_yrs, country, gender, birth_date, nationality, cor_grade, data_year)) %>% 
  mutate(empl_status = "Active",
         birth_date = parse_date_time(birth_date, orders = c("dmy"))) 
  
write_csv(active_data_combined, "processed/active_os_amr_2019-2021.csv")

# promos
promotions_os_amr_2019_2021 <- promo_amr_OS %>%
  select(!c(company, tybe_of_emp, level_3:position)) %>%
  rename(promotion_date = adjustment_date)
  
write_csv(promotions_os_amr_2019_2021, "processed/promotions_os_amr_2019-2021.csv")

# attition
attrition_raw_fine <- attrition_amr_os %>%
  mutate(text_4= "Inactive" ) %>% select(-gender) %>% rename(gender= gender_tex) %>%
  select(empl_no,employee_name, text_4, position_title, job_family, hire_date, 
         serv_yrs, cor_grade, leave_date, reason_text, reason_type) %>% 
  mutate(hire_date = parse_date_time(hire_date, orders= "dmy"),
         service_years = round(interval(hire_date, leave_date)/years(1),1)) %>% 
  mutate(change = case_when(
    service_years < 0 ~ "yes",
    service_years >=0 ~ "no",
                TRUE ~ NA_character_
  )) %>% filter(change== "no")

attrition_raw_fix <- attrition_amr_os %>%
  mutate(text_4= "Inactive" ) %>% select(-gender) %>% rename(gender= gender_tex) %>%
  select(empl_no,employee_name, text_4, position_title, job_family, hire_date, 
         serv_yrs, cor_grade, leave_date, reason_text, reason_type) %>% 
  mutate(hire_date = parse_date_time(hire_date, orders= "dmy"),
         service_years = round(interval(hire_date, leave_date)/years(1),1)) %>% 
  mutate(change = case_when(
    service_years <0 ~ "yes",
    service_years >=0 ~ "no",
                TRUE ~ NA_character_
  )) %>% filter(change == "yes" | is.na(change)) # need to include NA_character [continue here]

attrition_raw_fixed <- attrition_raw_fix %>%
  rename(hire_date = leave_date,
         leave_date = hire_date) %>% 
 mutate(service_years = round(interval(hire_date, leave_date)/years(1),1)) 

attrition_os_amr_2019_2021 <- bind_rows(
  attrition_raw_fine %>% select(-change),
  attrition_raw_fixed %>% select(-change)) %>%
  
  rename(status= text_4) %>%
  relocate(leave_date, .after=cor_grade) %>%
  relocate(service_years, .before=reason_text) %>%
  relocate(reason_type,.before=reason_text) %>%
  
  select(-serv_yrs) %>% filter(reason_text != "System Transfer") 


write_csv(attrition_os_amr_2019_2021, "processed/attrition_os_amr_2019-2021.csv")

# Tabular form data.

tbl <- left_join(active_data_combined %>% count(data_year, name= "active") ,
          attrition_os_amr_2019_2021 %>% count(year(leave_date), name= "attrition") %>% rename(data_year = 'year(leave_date)'))

tbl_all <- left_join(tbl,
                     promotions_os_amr_2019_2021 %>% count(year, name= "promotions") %>% rename(data_year = year)) %>%
  mutate(remarks = c("full_year", "full_year", "1 Jan to 30 September"))

write_csv(tbl_all, "processed/Tabular_data_OS_AMR_2019-2021.csv")

```

# Data Exploration

```{r}

# has the ae profile changed oiver time

active_data_combined %>% 
  mutate(data_year = as_factor(data_year)) %>%
  ggplot(aes(serv_yrs, data_year))+
geom_density_ridges(
    jittered_points = TRUE, quantile_lines = TRUE, scale = 0.7, alpha = 0.7,
    vline_size = 2, vline_color = "red",
    point_size = 0.4, point_alpha = 1,
    position = position_raincloud(adjust_vlines = TRUE)
  )+
  scale_x_continuous(name= "\nService years", breaks = seq(0,60,2))+
  scale_y_discrete(name="")+
  labs(title = "Distribution of service years is the narrowest in 2021",
       subtitle = "\nThe bottom quartile experience is at 3 years because we have not been able to hire new workers!\nAnd the top quartile experience is at 27 years because we have lost experienced workforce.\nIts a double whammy we are losing experienced workforce and and are unable to hire new workers.")

```
